\thetitle

### 2021-03-02 writing/challenges.tex
a bit more work on challenges
small tweaks to emphasize interesting mutations

However, three major challenges remain:  comparing the effect of
blame on the rational programmer across three different mechanisms; the
collection of a large number of representative debugging scenarios;  and a
proper sampling strategy for the huge space of possibilities. A
coincidental challenge is the disparity of the implementations of
gradually typed languages. To eliminate this variable from the comparison,
this paper uses Racket, which is thus far the only language in which all
three major soundness mechanisms are available in a reasonably robust and
directly comparable manner; for the implementation of the transient
semantics, see the parallel submission of an experience
report~\cite{ttt21}.

@@

Besides the run-time checks, pure luck might also amplify the usefulness
of blame. Thus a proper evaluation of blame needs to factor out any random
effects. A random rational programmer mode provides yet another necessary
baseline using the same ingredients as for exceptions.

The {\em second major challenge\/} is to find a representative collection of
programs with impedance mismatches. The latter must represent mistakes that programmers
accidentally create and that the run-time checks of academic systems catch. In
other words, the experiment calls for a collection of type mistakes in
mix-typed programs that is representative of those ``in the wild.''
Unfortunately no such collection exists, and with good reason. The kind of
mistakes needed are typically detected by unit or integration tests; even if it
takes some time to find their sources, these mistakes do not make it into code
repositories with appropriate commit messages.

@@

The {\em third major challenge\/} is the explosive number of debugging scenarios that
result from the combination of mutation-based scenario generation and mode-based
analysis. All three factors---three different gradual typing systems, the large
number of mutants, and the number of debugging modes---contribute possibly
useful experimental data in a multiplicative manner. Hence, carried out naively,
the experiment would demand an unacceptably large amount of computational
resources.  A practical execution has no option but to {\em sample the space of
scenarios\/}, carefully ensuring that it still produces reproducible
conclusions.

>>>>

However, three challenges remain:  comparing the effect of
blame on the rational programmer across three different mechanisms; the
collection of a large number of representative debugging scenarios;  and a
proper sampling strategy for the huge space of possibilities. A
coincidental challenge is the disparity of the implementations of
gradually typed languages. To eliminate this variable from the comparison,
this paper uses Racket, which is thus far the only language in which all
three major soundness mechanisms are available in a reasonably robust and
directly comparable manner; for the implementation of the transient
semantics, see the parallel submission of an experience
report~\cite{ttt21}.

@@

Besides the run-time checks, pure luck might also amplify the usefulness
of blame. Thus a proper evaluation of blame needs to factor out any random
effects too. A random rational programmer mode provides yet another necessary
baseline using the same ingredients as for exceptions.

The {\em second challenge\/} is to find a representative collection of
programs with impedance mismatches. The latter must represent mistakes that programmers
accidentally create and that the run-time checks of academic systems catch. In
other words, the experiment calls for a collection of type mistakes in
mix-typed programs that is representative of those ``in the wild.''
Unfortunately no such collection exists, and with good reason. The kind of
mistakes needed are typically detected by unit or integration tests; even if it
takes some time to find their sources, these mistakes do not make it into code
repositories with appropriate commit messages.

@@

The {\em third challenge\/} is the explosive number of debugging scenarios that
result from the combination of mutation-based scenario generation and mode-based
analysis. All three factors---three different gradual typing systems, the large
number of mutants, and the number of debugging modes---contribute possibly
useful experimental data in a multiplicative manner. Hence, carried out naively,
the experiment would demand an unacceptably large amount of computational
resources.  A practical execution has no option but to {\em sample the space of
scenarios\/}, carefully ensuring that it still produces reproducible
conclusions.
